Rectangulotopes

### 2024-04-02 draft/rectangulotopes.tex
minor changes

% math commands
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % set notation
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % big set notation
\newcommand{\Bigset}[2]{\Big\{ #1 \;\Big|\; #2 \Big\}} % Big set notation
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % set notation
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\dotprod}[2]{\left\langle \, #1 \; \middle| \; #2 \, \right\rangle} % dot product
\newcommand{\symdif}{\,\triangle\,} % symmetric difference
\newcommand{\one}{\b{1}} % the all one vector
\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\simplex}{\b{\triangle}} % simplex
\renewcommand{\implies}{\Rightarrow} % imply sign
\newcommand{\transpose}[1]{{#1}^t} % transpose matrix

@@

% polytopes
\newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron
% rectangulotopes
\newcommand{\WRP}{\polytope{WR}} % weak rectangulotope
\newcommand{\SRP}{\polytope{SR}} % strong rectangulotope
% shard polytopes
\newcommand{\SP}{\polytope{SP}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics. An elementary yet striking example of this idea is the \defn{permutahedron} $\polytope{P} (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$. The facets of $\polytope{P} (n)$ are indexed by the nonempty and nontrivial subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions. The many generalizations and extensions of the permutohedra gave rise to a flourishing theory of so-called \defn{generalized permutohedra}.

Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{MR4675114}. It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon. In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.

\begin{theorem}
  \label{thm:loday}
  Let $\Asso (n)$ be the convex hull of the points
  $\sum_{i\in [n]} w_i\cdot \mathbf{e}_i$, where
  \[
  w_i := \ell(T, i)\cdot r(T,i),
  \]
  for all binary trees $T$ on $n$ internal nodes, where $\ell(T,i)$ and $r(T,i)$ denote respectively the number of leaves in the left and right subtrees of $i$ in $T$.
  Then $\Asso (n)$ is a realization of the $(n-1)$-dimensional associahedron.
\end{theorem}

A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron. The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}. Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class. Those polytopes are generalized permutohedra that they called \defn{quotientopes}.

@@

The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with so-called \defn{strong rectangulations}~\cite{MR2864445}. The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{ACFF24}.

One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Loday's realization in Theorem~\ref{thm:loday}.
The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$. Similarly, the polytopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
  The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (See also Asinowski et al.~\cite{ACFF24}).
Rectangulotopes enrich the family of natural quotientopes, alongside associahedra and permutreehedra~\cite{MR3856522}, and are instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.

@@

\begin{figure}
	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
	\caption{\label{fig:sourceTargetTrees}A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
\end{figure}

Fix a rectangulation~$R$ with $n$ rectangles, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right, and the vertical edges are oriented from bottom to top.
Hence, the bottom left corner of~$r$ is its source~$s(r)$, while its top right corner is its target~$t(r)$.
Note that since~$R$ is generic, the sources~$\set{s(r)}{r \in [n]}$ and the targets~$\set{t(r)}{r \in [n]}$ are disjoint.
The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in [n]}$ (resp.~by the targets~$\set{t(r)}{r \in [n]}$).
See \cref{fig:sourceTargetTrees}.

@@

We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:sourceTargetTrees}.
Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.

@@

\begin{theorem}
  Let $\WRP (n)$ be the convex hull of the points
  $\sum_{i\in [n]} (\loday{w}_i - \antiloday{w}_i)\cdot \mathbf{e}_i$,
  where
  \[
  \begin{split}
    \loday{w}_i := & h(T, i)\cdot v(T,i) \\
    \antiloday{w}_i := & h(S, i)\cdot v(S,i),
  \end{split}
  \]
  for all weak rectangulations $R$ on $n$ rectangles, where $T:=T(R)$, $S:=S(R)$,
  and $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
  Then $\WRP (n)$ is a realization of the $(n-1)$-dimensional weak rectangulotope.
\end{theorem}

@@

\begin{theorem}
  Let $\SRP (n)$ be the convex hull of the points
  \[
  \sum_{i,j\in [n], i< j} (\yin{w}_{i,j} - \yang{w}_{i,j})\cdot (\mathbf{e}_i - \mathbf{e}_j),
  \]
  with
  \[
  \begin{split}
    \yin{w}_{i,j} := & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
    \yang{w}_{i,j} := & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket
  \end{split}
  \]

@@

  % jean:{Double check the formula.}
  for all strong rectangulations $R$ on $n$ rectangles, where:
  \begin{itemize}
  \item $T := T(R)$ and $S:=S(R)$,
  \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of common leaves of the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
  \end{itemize}
Then $\SRP (n)$ is a realization of the $(n-1)$-dimensional strong rectangulotope.  
\end{theorem}

@@

@@

Given a rectangulation $R$, the two trees $T(R)$ and $S(R)$, where the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top,
define two partial orders on $[n]$.
We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$~\cite{MR4014603}. The weak poset is a two-dimensional lattice, the minimum of which is the root of $S(R)$ and the maximum is the root of $T(R)$.
We define the set $\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
\[
\mathcal{L}(P_w(R)) := \{\pi\in\f{S}_n : i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
\]

@@

\begin{figure}
	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right). The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.}
\end{figure}

@@

@@

\begin{lemma}
  \label{lem:lodaymax}
  Let $\pi\in\f{S}_n$ and consider a Loday arc $A$ defined by an interval $I$.
  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
  is $\mathbf{e}_i$, where $i := \arg\max_{j\in I} \pi^{-1}(j)$.
  Similarly, if $A$ is an anti-Loday arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
  is $-\mathbf{e}_i$, where $i := \arg\min_{j\in I} \pi^{-1}(j)$.
\end{lemma}

@@

We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$, and $\mathcal{L}(P_s)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P_s$:
\[
\mathcal{L}(P_s(R)) := \{\pi\in\f{S}_n : i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
\]

@@

\begin{lemma}
  \label{lem:yinminmax}
  Let $\pi\in\f{S}_n$ and consider a yin arc $A=(I,J)$.
  Let $i := \arg\max_{k\in I} \pi^{-1}(k)$ and $j := \arg\min_{k\in J} \pi^{-1}(k)$.
  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
  is $\mathbf{e}_i-\mathbf{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\mathbf{0}$ otherwise.
\end{lemma}

>>>>

% math commands
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % set notation
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % big set notation
\newcommand{\Bigset}[2]{\Big\{ #1 \;\Big|\; #2 \Big\}} % Big set notation
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % set notation
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\dotprod}[2]{\left\langle \, #1 \; \middle| \; #2 \, \right\rangle} % dot product
\newcommand{\symdif}{\,\triangle\,} % symmetric difference
\newcommand{\one}{\b{1}} % the all one vector
\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % \eqdef
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\simplex}{\b{\triangle}} % simplex
\renewcommand{\implies}{\Rightarrow} % imply sign
\newcommand{\transpose}[1]{{#1}^t} % transpose matrix

@@

% polytopes
\newcommand{\Perm}{\polytope{P}} % permutahedron
\newcommand{\Asso}{\polytope{A}} % associahedron
% rectangulotopes
\newcommand{\WRP}{\polytope{WR}} % weak rectangulotope
\newcommand{\SRP}{\polytope{SR}} % strong rectangulotope
% shard polytopes
\newcommand{\SP}{\polytope{SP}}

@@

\section{Introduction}

@@

The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics. An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$. Its facets are indexed by the proper nonempty subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions. The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra} (also known as \defn{generalized permutahedra}).

Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{MR4675114}. It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
The face lattice of the associahedron~$\Asso (n)$ is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon. In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.

\begin{theorem}
  \label{thm:loday}
  Let $\Asso (n)$ be the convex hull of the points
  $\sum_{i\in [n]} w_i\cdot \mathbf{e}_i$, where
  \[
  w_i \eqdef \ell(T, i)\cdot r(T,i),
  \]
  for all binary trees $T$ on $n$ internal nodes, where $\ell(T,i)$ and $r(T,i)$ denote respectively the number of leaves in the left and right subtrees of $i$ in $T$.
  Then $\Asso (n)$ is a realization of the $(n-1)$-dimensional associahedron.
\end{theorem}

A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron. The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}. Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class. Those polytopes are generalized permutahedra that they called \defn{quotientopes}.

@@

The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}. The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{ACFF24}.

One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Loday's realization in Theorem~\ref{thm:loday}.
The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
Similarly, the quotientopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
The skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (see also~\cite{ACFF24}).

Rectangulotopes enrich the family of natural quotientopes, alongside associahedra and permutreehedra~\cite{MR3856522}, and are instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.

@@

\begin{figure}
	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
	\caption{A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
        \label{fig:strongRectangulation}
\end{figure}

Fix a rectangulation~$R$ with $n$ rectangles, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right, and the vertical edges are oriented from bottom to top.
Hence, the bottom left corner of~$r$ is its source~$s(r)$, while its top right corner is its target~$t(r)$.
Note that since~$R$ is generic, the sources~$\set{s(r)}{r \in [n]}$ and the targets~$\set{t(r)}{r \in [n]}$ are disjoint.
The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in [n]}$ (resp.~by the targets~$\set{t(r)}{r \in [n]}$).
See \cref{fig:strongRectangulation}.

@@

We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:strongRectangulation}.
Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.

@@

\begin{theorem}
  Let $\WRP (n)$ be the convex hull of the points
  $\sum_{i\in [n]} (\loday{w}_i - \antiloday{w}_i)\cdot \mathbf{e}_i$,
  where
  \[
  \begin{split}
    \loday{w}_i \eqdef & h(T, i)\cdot v(T,i) \\
    \antiloday{w}_i \eqdef & h(S, i)\cdot v(S,i),
  \end{split}
  \]
  for all weak rectangulations $R$ on $n$ rectangles, where $T \eqdef T(R)$, $S \eqdef S(R)$,
  and $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
  Then $\WRP (n)$ is a realization of the $(n-1)$-dimensional weak rectangulotope.
\end{theorem}

\begin{example}
The vertex of~$\WRP (16)$ corresponding to the weak rectangulation of \cref{fig:weakRectangulation} is
\[
(-3, 0, 26, 2, -9, 5, -69, -4, 17, 0, -8, 59, 2, -20, 0, 2).
\]
For instance, the $4$th coordinate is $h(T, 4) \cdot v(T, 4) - h(S, 4) \cdot v(S, 4) = 1 \cdot 5 - 1 \cdot 3 = 2$.
\end{example}

@@

\begin{theorem}
  Let $\SRP (n)$ be the convex hull of the points
  \[
  \sum_{i,j\in [n], i< j} (\yin{w}_{i,j} - \yang{w}_{i,j})\cdot (\mathbf{e}_i - \mathbf{e}_j),
  \]
  with
  \[
  \begin{split}
    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket
  \end{split}
  \]

@@

  % jean:{Double check the formula.}
  for all strong rectangulations $R$ on $n$ rectangles, where:
  \begin{itemize}
  \item $T \eqdef T(R)$ and $S \eqdef S(R)$,
  \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of common leaves of the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
  \end{itemize}
Then $\SRP (n)$ is a realization of the $(n-1)$-dimensional strong rectangulotope.  
\end{theorem}

@@

\begin{example}
The vertex of~$\WRP (16)$ corresponding to the weak rectangulation of \cref{fig:strongRectangulation} is
\[
???
\]
For instance, the $4$th coordinate is $???$.

\end{example}

@@

Given a rectangulation $R$, the two trees $T(R)$ and $S(R)$, where the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top,
define two partial orders on $[n]$.
We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$~\cite{MR4014603}. The weak poset is a two-dimensional lattice, the minimum of which is the root of $S(R)$ and the maximum is the root of $T(R)$.
We define the set $\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
\[
\mathcal{L}(P_w(R)) \eqdef \{\pi\in\f{S}_n : i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
\]

@@

\begin{figure}
	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right).}
	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
	\label{fig:weakRectangulation}
\end{figure}

@@

\vincent{Say also the minkowski sum of all faces of the standard simplex corresponding to intervals, and all their opposite.}

@@

\begin{lemma}
  \label{lem:lodaymax}
  Let $\pi\in\f{S}_n$ and consider a Loday arc $A$ defined by an interval $I$.
  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
  is $\mathbf{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$.
  Similarly, if $A$ is an anti-Loday arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
  is $-\mathbf{e}_i$, where $i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$.
\end{lemma}

@@

We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$, and $\mathcal{L}(P_s)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P_s$:
\[
\mathcal{L}(P_s(R)) \eqdef \{\pi\in\f{S}_n : i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
\]

@@

\begin{lemma}
  \label{lem:yinminmax}
  Let $\pi\in\f{S}_n$ and consider a yin arc $A=(I,J)$.
  Let $i \eqdef \arg\max_{k\in I} \pi^{-1}(k)$ and $j \eqdef \arg\min_{k\in J} \pi^{-1}(k)$.
  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
  is $\mathbf{e}_i-\mathbf{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\mathbf{0}$ otherwise.
\end{lemma}
